---
title: "Reproducible Research"
author: "Ryan Hammer"
date: "3/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Welcome to a set of user created course notes for Reproducible Research, the 5th of 10 courses in the Johns Hopkins Data Specialization available through Coursera. This document is meant to serve as a study guide in conjunction with watching the instructional videos and attempting the optional Swirl exercises.
For addional resources, visit the community forum on the course page on Coursera, as well as the link below:



Without further ado, we begin.

## Reproducible Research Best Practices

-Start from raw data and include it through a link or source file in the final output. Then include code for any cleaning of the data that is done, rather than just starting from pre-processed data.
-Use version control to help keep track of changes being made
-Use statistical software whose operation can be coded; this allows for creation of a set of instructions for how the data has been manipulated.
-Save data in non-propietary formats.

## Knitr

Knitr is a form of literate statistical programming. Literate statistical programming solves the problem of combining data, code, and analysis into one document. It creates an article that is a stream of text and code, with each broken into chunks. The result is a document that is easy to read for humans in its output format, but also contains code that is machine readable and executable as well.
This allows for more easily reproducible research because readers can view the code used to take the raw data and manipulate it for use in whatever analysis is being done; more simply put, it creates the pudding containing the proof. Any person lending a critical eye to a set of analysis can copy and paste the code and reproduce the results on the raw data (or not if the proverbial pudding is bad).
---
Knitr is an R package available on CRAN and comes integrated into R Studio. It supports R Markdown, LaTeX, or HTML and can export to PDF or HTML. It is extremely useful for creating manuals, technical documents, tutorials, reports, and/or data preprocessing documents and summaries.

## Text Formatting in R Markdown

To italicize text in R Markdown, use a single * at the beginning and end of the word or phrase for which italics are desired. For example, to italicize "Reproducible Research" would be as follows:
```{r}

```



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
